---
editor_options: 
  chunk_output_type: console
---

# Making datasets comparable

In this script, we will ensure that the point count data and the acoustic data are comparable to one another. We apply a series of filters to ensure that we account for broader differences in sampling methodologies to ensure that they are comparable to one another. 

## Install required libraries  

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(sf)
library(raster)

# for plotting
library(scales)
library(ggplot2)
library(ggspatial)
library(colorspace)

# Source any custom/other internal functions necessary for analysis
source("code/01_internal-functions.R")
```

## Loading point count data

```{r}
# This is data from Hariharan and Raman (2021), which consists of point counts carried out across 69 sites
point_counts <- read.csv("data/point-count-data.csv")

# add eBird species code to the point-count data for comparison with acoustic data later on
species_codes <- read.csv("data/species-annotation-codes.csv")

# for the sake of comparability, some issues were fixed in the point-count dataset
# please note that these issues were fixed using the .csv file and not through R
# the following issues were fixed:
# leading spaces were removed from the common_name
# species that start with the common_name 'Grey' were changed to 'Gray'

# merge with point_count data
point_counts <- left_join(point_counts, species_codes, by = "common_name")

# removing all mammal species and unidentified bird species
point_counts <- point_counts %>%
  filter(bird_mammal == "Bird")
```

## Loading the acoustic data

The acoustic data consists of annotations of 10-s chunks of audio files across summer and winter months. We will load both these annotation sets and process them further to ensure that it is comparable to the point count dataset.  
```{r}
# Attach the annotation data for summer and winter
# These two .csvs below are not uploaded to GitHub and can be provided upon request
# Please write to Vijay Ramesh if you would like to access the raw data - vr292@cornell.edu)

summer_data <- read.csv("data/summer-dawn-annotations.csv")
winter_data <- read.csv("data/winter-dawn-annotations.csv")

# combine the datasets to a single dataframe
acoustic_data <- bind_rows(summer_data, winter_data)
names(acoustic_data)

# reorder columns to ensure the species codes appear one after another
# also include species that are present in point count data to later combine both dataframes together
acoustic_data <- acoustic_data %>%
  relocate(c("BFOW", "SBEO", "JUNI", "ASKO", "HSWO", "TBWA"), .after = "CORO")

# split the filename column into 4 columns : Site, Date, Time and Splits
acoustic_data <- separate(acoustic_data, col = Filename, into = c("site", "date", "time", "splits"), sep = "_")
```


## Subset acoustic data and point count data

To ensure that the datasets are comparable, we carry out the following:

a)  Subset only data from those sites for which both point count data and acoustic data was collected.

b)  Ensure that the dates of visits (for both point count data and acoustic data) are coming from the similar seasons, if not the same months.

It is important to clarify that while the point count data and acoustic data were collected across similar months (to match seasons), majority of the point count data were collected between November 2019 and March 2020, while the majority of the acoustic data were collected in March 2020, a few days in May 2020 (covid-19 related delays) and between December 2020 and January 2021.

c)  Similar levels of effort across sites (for point count data and acoustic data). Here, we define effort as the total number of minutes a site was surveyed. The duration of each point count at a site was 15 minutes long while the duration of each acoustic survey at a site was 16 minutes in duration. For the sake of comparability, we will compare 15-minutes of point count data to 15-minutes of acoustic data (by removing 1-min from each of acoustic 'visits'). This will result in a total of 90 minutes of effort at each site across point count and acoustic data.   

```{r}
# a subset of sites that were included in the Hariharan and Raman (2021) study were included in the Ramesh et al. (2023) study due to logistical constraints of placing audio recorders
# we exclude OLCAP5B from the audio data as very few acoustic visits were carried out at this site
sites <- read.csv("data/list-of-sites.csv") %>%
  filter(site_id != "OLCAP5B")

# convert date column to YMD format using lubridate::ymd()
point_counts$date <- lubridate::ymd(point_counts$date)

# subset point count data to only include the above list of sites
point_counts <- point_counts %>%
  mutate(site_id = str_replace_all(site_id, "_", ""))

point_counts <- left_join(sites[, c(2, 3, 4, 5)],
  point_counts,
  by = c("site_id" = "site_id")
)

# To ensure that the datasets are comparable, we will choose a maximum of six visits to each site from the point count study (translating to an effort of 90 minutes per site)
nSitesDays <- point_counts %>%
  dplyr::select(site_id, date) %>%
  distinct() %>%
  arrange(site_id) %>%
  count(site_id)

# we observed that some sites had as many as 7 visits. We will subset these to choose only six visits (see reasoning above)

# unique date site combination to give you a sense of sampling
uniqueSiteDate <- point_counts %>%
  group_by(site_id) %>%
  distinct(date)

# the below lines of code were written following a query on stackOverflow to select six non-consecutive visits to any site
# Link: https://stackoverflow.com/questions/67212152/select-non-consecutive-dates-for-every-grouped-element-in-r

nonConVisits <- uniqueSiteDate %>%
  ungroup() %>%
  group_split(site_id) %>%
  map_df(., ~ .x %>%
    ungroup() %>%
    arrange(date) %>%
    mutate(n = 1) %>%
    complete(date = seq.Date(first(date), last(date),
      by = "days"
    )) %>%
    group_by(n = cumsum(is.na(n))) %>%
    filter(!is.na(site_id)) %>%
    filter(row_number() %% 2 == 1) %>%
    ungroup() %>%
    sample_n(min(n(), 6))) %>% # change n here for nvisits
  dplyr::select(-n)

# left-join with the original dataframe to subset the data
pc_subset <- left_join(nonConVisits, point_counts)

# subset acoustic data (similar to point count data)

# first we will remove OLCAP5B - a site for which only 3 visits were made in summer and not sampled as a result of logistic reasons in winter
acoustic_data <- acoustic_data %>%
  filter(!str_detect(site, "OLCAP5B"))

# Convert date column to YMD format using lubridate::ymd()
acoustic_data$date <- lubridate::ymd(acoustic_data$date)

# We do not exclude data from May 2020 for the acoustic dataset any longer as this will result in a reduction in sample size
# acoustic_data <- acoustic_data %>%
#  filter(!(date >= "2020-05-01" & date <= "2020-05-10"))

# number of visits to a particular site
# some sites have as many as 7 visits, but many visits have only 5 visits and only INBS04U has 4 visits
# for the sake of comparability with point count data, we choose 5 visits across sites translating to effort of 80 min per site
nSitesDays <- acoustic_data %>%
  dplyr::select(site, date) %>%
  distinct() %>%
  arrange(site) %>%
  count(site)

# unique date site combination to give you a sense of sampling
uniqueSiteDate <- acoustic_data %>%
  group_by(site) %>%
  distinct(date)

# note: We could not choose non-consecutive days for the acoustic data due to heavy rain and we could only sample consecutive days for some sites
# for the acoustic data, we choose 6 random visits per site
randVisits <- uniqueSiteDate %>%
  ungroup() %>%
  group_split(site) %>%
  map_df(
    ., ~ .x %>%
      ungroup() %>%
      arrange(date) %>%
      mutate(n = 1) %>%
      complete(date = seq.Date(first(date), last(date),
        by = "days"
      )) %>%
      group_by(n = cumsum(is.na(n))) %>%
      filter(!is.na(site)) %>%
      ungroup() %>%
      sample_n(min(n(), 6)) # change n here for number of visits
  ) %>%
  dplyr::select(-n)

# left-join with the original dataframe to subset the data for analysis
aru_subset <- left_join(randVisits, acoustic_data)

# remove the last 1-minute of data from each acoustic 'visit'
# since the data is split into 10-s chunks, we want to remove the last 60-seconds from the fourth 4-min segment of each acoustic visit 

# create row numbers for every distinct 4-minute segment within a single 16-minute visit
# this is done so that we can essentially get rid of the last minute from the fourth segment. 
add_row_number <- aru_subset %>%
  dplyr::select(site, date, time) %>%
  distinct() %>%
  group_by(site, date) %>%
  mutate(instance = row_number())

# add this back to the previous subset of data
aru_subset <- left_join(aru_subset, add_row_number)

# remove the last-minute of data from each visit
# these are the 10-s second splits that correspond to 60 seconds of data for the last minute
target <- c("180-190","190-200", "200-210",
            "210-220","220-230","230-240")

# let's filter out the data using the row_number (4, which translates to the fourth segment of the 16-minutes)
aru_to_remove <- aru_subset %>%
  filter(instance == 4 & !splits %in% target)

# prior to joining both datasets
# get rid of the fourth instance of data
for_aru_subset <- aru_subset %>%
  filter(!instance == 4)

aru_subset <- full_join(for_aru_subset, aru_to_remove)
```

## Evaluating both subsets of data before combining them into a single dataframe for further analysis

For the point count dataset, we shall remove the extra columns. For the acoustic dataset, we will dplyr::pivot_longer() and add the same columns as point count dataset before binding rows to create a single dataframe.  
```{r}
# remove unnecessary columns from the point-count dataset
pc_subset <- pc_subset[, -c(4, 5, 6, 9, 11, 15, 17, 18)]

# add a time of day column to the point-count dataset for future calculations and a data_type column
pc_subset <- pc_subset %>%
  mutate(
    time_of_day =
      case_when(
        start_time >= 6 & start_time < 7 ~ "6AM to 7AM",
        start_time >= 7 & start_time < 8 ~ "7AM to 8AM",
        start_time >= 8 & start_time < 9 ~ "8AM to 9AM",
        start_time >= 9 & start_time <= 10 ~ "9AM to 10AM"
      )
  ) %>%
  mutate(data_type = "point_count")

# rename acoustic data columns
aru_subset <- rename(aru_subset, site_id = site)
aru_subset <- rename(aru_subset, restoration_type = Restoration.Type..Benchmark.Active.Passive.)

# pivot_longer the acoustic data and remove zero values
aru_subset <- aru_subset %>%
  group_by(site_id, date, time) %>%
  transform() %>%
  replace(is.na(.), 0) %>%
  summarise_at(.vars = vars(c("INPE":"TBWA")), .funs = sum) %>%
  pivot_longer(cols = INPE:TBWA, names_to = "eBird_codes", 
               values_to = "number") %>%
  filter(number != 0) # one way to remove zeros

# add a start_time column based on the time the acoustic visit was started at that site-day combination
# for example, if HP36P1B's first data point/visit came from 091000, then the start_time for all the 15-min would have the value 091000
aru_subset <- aru_subset %>%
  group_by(site_id, date) %>%
  mutate(start_time = min(time)) %>%
  ungroup()

# add time_of_day column to indicate time-windows for when the acoustic-visit was started and add a data_type column
aru_subset <- aru_subset %>%
  mutate(
    time_of_day =
      case_when(
        start_time >= "060000" & start_time < "070000"
        ~ "6AM to 7AM",
        start_time >= "070000" & start_time < "080000"
        ~ "7AM to 8AM",
        start_time >= "080000" & start_time < "090000"
        ~ "8AM to 9AM",
        start_time >= "090000" & start_time <= "100000"
        ~ "9AM to 10AM"
      )
  ) %>%
  mutate(data_type = "acoustic_data")

# rename columns and add additional columns to ensure it is comparable to the point count dataset
aru_subset <- rename(aru_subset, time_segment = time)
aru_subset$HSF <- "H"
aru_subset$distance <- NA

names(aru_subset)[4] <- "species_annotation_codes"

# add species scientific_name and common_name to the aru_subset data
aru_subset <- left_join(aru_subset, species_codes[, c(1, 2, 3)],
  by = "species_annotation_codes"
)
```

## Bind the acoustic and point count datasets into a single dataframe

The subsequent dataframe from this code chunk can be used directly in the following scripts to run other analyses.  
```{r}
# check the structure and names of both subsets prior to creating a single dataframe
str(pc_subset)
str(aru_subset)

# change structures to match across subsets
pc_subset$start_time <- as.character(pc_subset$start_time)
pc_subset$time_segment <- as.character(pc_subset$time_segment)

# bind_rows to create a single dataframe
datSubset <- bind_rows(pc_subset[,-c(3,11)], aru_subset[,-4])

# write to file
write.csv(datSubset, "results/datSubset.csv", row.names = F)
```

**Metadata for the above dataframe is provided below:**

**date**: Date of the point count/acoustic survey (str: Date)

**siteid**: Site name and can be cross-referenced to the *list-of-sites.csv* file (str: Character)

**restorationType**: Sites can be actively restored (AR), passively restored (naturally regenerating/NR) or benchmark (undisturbed/BM) forests (str: Character)

**startTime**: Start time for the point count and the acoustic data (str: Character)

**timeSegment**: The point count data was carried out in three 5-min segments and indicated as 1, 2, and 3. The acoustic data was segmented into four 4-min segments and indicated by the startTime of each segment unlike the point count data (str: Character)

**commonName**: Species common name (str: Character)

**number**: For the point count data, this column indicates the number of individuals of a species seen/heard/flying above within a time segment. For the acoustic data, this column indicates the number of vocalizations of a species within a time segment (str: Integer)

**distance**: For the point count data, this column indicates the distance at which an individual of a species was observed (seen) within a time segment. This number varied from 0 to 50 metres and all individuals/species above that distance was not reported in the point count dataset. For the acoustic dataset, such values are not available and marked as NA (str: Integer)

**HSF**: Heard/Seen/Flying was recorded for the point count dataset and indicated by H/S/F. For the acoustic data, the column has been filled with the letter H only (str: Character)

**scientificName**: Species scientific name (str: Character)

**timeOfDay**: Depending on the start times of the point count or the acoustic data survey/visit, this column indicates if a survey was within any one-hour duration between 6AM to 10AM. For example, if the start time of the point count was '7.42', this column would indicate the timeOfDay as '7AM to 8AM'. Similarly, if the start time of the acoustic survey was '091000', this column would indicated the timeOfDay as '9AM to 10AM' (str: Character)

**dataType**: Point count or acoustic data (str: Character)
